---
title: "Salmonellen NGS-Cluster/Ausbruch Zusammenfassung"
date: "`r format(Sys.Date(), format = '<br/>Berichtsstand %d.%m.%Y')`"
lang: de-DE
params:
  author_name:
    label: "Name des Autors:"
    value: "Ihr Name"
    input: text
  author_affiliation:
    label: "Zugehörigkeit des Autors:"
    value: "Abteilung für Infektionsepidemiologie, Robert Koch-Institut"
    input: text
  outbreak_declared:
    label: "Wurde dieser Cluster zu einem Ausbruch erklärt?"
    value: TRUE
    input: checkbox
  previous_report:
    label: "Vergleich mit früherem Bericht?"
    value: FALSE
    input: checkbox
  example_data:
    label: "Testbericht mit Beispieldaten?"
    value: TRUE
    input: checkbox
  data_password: 
    label: "Datenkennwort:"
    value: ""
    input: password
  data_current:
    label: "Aktuelle Zeilenliste auswählen:"
    value: "data/ll_current.rds.enc"
    input: file
  data_previous:
    label: "Vorherige Zeilenliste auswählenn:"
    value: "data/ll_previous.rds.enc"
    input: file
  map_breaks:
    label: "Wählen Sie die Bruchmethode für die Choroplethenkarte:"
    value: fisher
    input: select
    choices: [fisher, geom, jenks, quantile, sd, pretty]
  lag_casedate:
    label: "Verzögerung (Wochen) für das Datum der Fallmeldung:"
    value: 1
    input: numeric
  lag_labdate:
    label: "Verzögerung (Wochen) für das Eingangsdatum der Probe:"
    value: 1
    input: numeric
  stratifier:
    label: "Stratifikator für Epikurven auswählen:"
    value: Bundesland
    input: radio
    choices: [Bundesland, Falldefinition]
output: 
  officedown::rdocx_document:
    reference_docx: "resources/vorlagendokument.docx"
    pandoc_args:
      - '--lua-filter=merge-code-blocks.lua'
    page_margins:
      bottom: 2.5
      top: 2.5
      right: 1.5
      left: 1.5
      header: 1
      footer: 1
      gutter: 0.5
    tables:
      style: Table
      layout: autofit
      width: 1.0
      caption:
       style: Table Caption
       pre: 'Tabelle '
       sep: ': '
    plots:
      style: Figure
      align: left
      topcaption: FALSE
      caption:
       style: Image Caption
       pre: 'Abb. '
       sep: ': '
urlcolor: blue
always_allow_html: yes
---

```{r setup, include=FALSE, purl=FALSE}

### Store width of page between margins in Microsoft Word template:

# Check width of page by opening vorlagendokument.docx and switching the ruler on
# Width of page between the margins in the word document (in centimeters)
pagewidth_cm <- 18

# Width of page converted to inches:
pagewidth_in <- pagewidth_cm * 0.3937 


# Set chunk options:
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE, 
                      eval = TRUE,
                      ft.align = "left",
                      fig.width = pagewidth_in,
                      fig.height = 4.5,
                      out.width = "100%")

# Check if pacman is installed, install if not:
if (!require("pacman")) install.packages("pacman")

# Install packages required to build the r markdown report:
pacman::p_load(
  rmarkdown,
  knitr, 
  shiny,
  bookdown,
  officedown,
  officer
)

# Add option to conditionally show text and evaluate inline expressions:
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})


```



```{r load_libraries}


# Use pacman::p_load_gh to load required GitHub packages:
pacman::p_load_gh(

  # Works with sf package to improve positioning of labels on maps:
  "yutannihilation/ggsflabel"

  )

# Use pacman::p_load to load required CRAN packages:
pacman::p_load(
  
  # Packages that make importing and exporting files and objects easier:
  here, 
  rio,
  safer,
  janitor,
  
  # Packages for creating nicely formatted summary tables:
  flextable,
  ftExtra,
  ztable,
  gtsummary, 
  
  # Add-on functions for ggplot graphics
  gridExtra,
  ggrepel,
  ggtext,
  scales, 
  
  # Packages for making geographic maps:
  sf, 
  mapsf,
  ggmap,
  
  # Packages containing epidemiological functions:
  epikit, 
  apyramid, 
  
  # Loads suite of essential 'tidyverse' packages including: 
  # dplyr, tidyr, lubridate, stringr, ggplot2, forcats, purrr, tibble & readr
  tidyverse

  )


```


```{r load_functions}

# For now the functions developed for this report are sourced individually here.
# Later they will be loaded as part of the package that contains the shiny app.

# Source function to calculate the year and week from a date:
source(here("functions", "get_yearweek.R"))

# Source function to calculate the year and month from a date:
source(here("functions", "get_yearmonth.R"))

# Source function to convert tables of numeric values to heatmap tables:
source(here("functions", "heatmaptable.R"))

# Source function to make calendar weeks or months human readable:
source(here("functions", "print_dateunit.R"))

# Source function to clean raw line list data:
source(here("functions", "clean_data.R"))

# Source function to calculate numbers and statements for executive summary:
source(here("functions", "execsum.R"))

# Source function to compare bundesland numbers and statements:
source(here("functions", "compare_bundesland.R"))

# Source function to compare values from old and new line lists:
source(here("functions", "compare_table.R"))

# Source function to emphasize text with blue colour and bold font:
source(here("functions", "colourtext.R"))

# Source function to calculate and label incidence breaks for map:
source(here("functions", "ibreak_labeller.R"))

# Source function to format epicurve with a double date x axis:
source(here("functions", "label_date_axis.R"))

# Source function to format bottom axis of age-sex pyramid with whole integers:
source(here("functions", "integer_breaks.R"))

```


```{r set_language}

# Set the language to German:
invisible(capture.output(Sys.setlocale("LC_ALL", "de_DE.UTF-8")))

### Note: wrapped in invisible(capture.output()) to suppress info message.
### Suppressing messages does not work for this as it is categorized as info.

```


```{r font_styles}

# Set flextable default fonts:
set_flextable_defaults(font.family = "Calibri", 
                       cs.family = "Calibri", 
                       hansi.family = "Calibri", 
                       eastasia.family = "Calibri", 
                       decimal.mark = ",", 
                       big.mark = ".")


# Create normal paragraph style:
nstyle <- officer::fp_text(color = "black", 
                           bold = FALSE, 
                           font.size = 10, 
                           font.family = "Calibri")

# Create highlight text style:
hstyle <- officer::fp_text(color = "#4F81BD", 
                           bold = TRUE, 
                           font.size = 10, 
                           font.family = "Calibri", 
                           shading.color = "#C6D9F1")

# Create a style for section headers that should not appear in the TOC:
nontoc_header <- officer::fp_text(
  color = "#4F81BD", 
  font.size = 16, 
  font.family = "Calibri",
  bold = TRUE
)

# Create style for summary text sub-headings:
abstract_header <- officer::fp_text(
  color = "black", 
  font.size = 12,
  font.family = "Calibri",
  bold = TRUE
)

# Create a style for placehoders for where to add manual text:
placeholder <- officer::fp_text(
  color = "darkgrey",
  font.size = 10,
  font.family = "Calibri"
)

# Set flextable defaults:
set_flextable_defaults(
  font.family = "Calibri", 
  font.size = 12
  )

```



```{r locate_linelists}

if (params$example_data) {
  
  # Create temporary file placeholder to hold raw bytes:
  tmp_current <- tempfile(fileext = ".xlsx")
  tmp_previous <- tempfile(fileext = ".xlsx")
  
  # Decrypt current line list file into raw byttes in R memory:
  raw_current <- safer::decrypt_file(
    infile = params$data_current, 
    key = params$data_password,
    outfile = tmp_current
    )
  
  # Decrypt previous line list file into raw bytes in R memory:
  raw_previous <- safer::decrypt_file(
    infile = params$data_previous,
    key = params$data_password,
    outfile = tmp_previous
    )

} else {
  
  # Otherwise just pass the relative file paths to location names:
  tmp_current <- params$data_current
  tmp_previous <- params$data_previous
  
}


```



```{r import_colour_codes}

# Import the colour codes to be applied to federal states (Bundesland):
farbcode <- rio::import(here("resources", "BL Farbcode.xlsx")) %>% 
  
  # Add the Hex codes for each colour in a new column:
  mutate(hexcode = rgb(red = R, 
                       green = G, 
                       blue = B,
                       maxColorValue = 255))

# Create key to check that assigned colours are correct:
farbcode_key <- ggplot(farbcode, aes(x = Bundesland, 
                                     y = 1, 
                                     fill = hexcode)) +
  geom_tile() +
  geom_text(aes(label = Code)) +
  scale_fill_identity() +
  coord_flip() +
  scale_x_discrete(limits = rev) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(), 
        axis.ticks.x = element_blank())

```


```{r import_population}

#########################################################
# Import population data for federal states (bundesland):

# Import data:
pop_bundesland <- rio::import(here("resources",
                                   "SBD_pop_12411_0010_DLAND_flat.csv"),
                              encoding = "Latin-1") %>% 
  
  # Clean up column names:
  janitor::clean_names() %>% 
  
  # Rename columns:
  rename(pop_date = "time", 
         admin_level_name = "x1_variable_code", 
         admin_level_type = "x1_variable_label", 
         bundesland_numcode = "x1_variable_code_2", 
         bundesland_name = "x1_variable_code_3", 
         bundesland_pop = "bevstd_population_number") %>% 
  
  # Filter for latest population figures only:
  filter(pop_date == max(pop_date, na.rm = TRUE)) %>% 
  
  # Pad bundesland code with leading 0s so that all are 2 characters long:
  mutate(bundesland_numcode = stringr::str_pad(string = bundesland_numcode, 
                                               width = 2, 
                                               side = "left", 
                                               pad = "0")) %>% 
  
  # Left-join the two-letter codes from farbcode:
  left_join(farbcode %>% select(Bundesland, Code, hexcode), 
            by = join_by(bundesland_name == Bundesland)) %>% 
  
  # Rename column with two-letter code:
  rename(bundesland_lettercode = "Code") %>% 


  # Select only necessary columns:
  select(bundesland_numcode,
         bundesland_lettercode,
         bundesland_name,
         hexcode,
         bundesland_pop)

  
#########################################################
# Import population data for counties (kreise):

# Import data:
pop_kreise <- rio::import(here("resources", 
                               "SBD_pop_12411_0015_KREISE_flat.csv"), 
                          encoding = "Latin-1") %>% 
  
  # Clean up column names:
  janitor::clean_names() %>% 
  
  # Rename columns:
  rename(pop_date = "time", 
         admin_level_name = "x1_variable_code", 
         admin_level_type = "x1_variable_label", 
         kreise_code = "x1_variable_code_2", 
         kreise_name_type = "x1_variable_code_3", 
         kreise_pop = "bevstd_population_number") %>% 
  
  # Filter for latest population figures only:
  filter(pop_date == max(pop_date, na.rm = TRUE)) %>% 
  
  # Separate district name and type into two columns:
  separate_wider_delim(cols = kreise_name_type, 
                       delim = ", ", 
                       names = c("kreise_name", "kreise_type"), 
                       too_few = "align_start") %>% 
  
  # Replace - with NA:
  mutate(kreise_pop = na_if(kreise_pop, "-")) %>% 
  
  # Filter out old divisions that are no longer included 
  # (Marked by expiry date that starts with 'until(date)' or 'b.date'):
  filter(grepl("\\(until|\\(b.", x = kreise_name) == FALSE 
         & grepl("\\(until|\\(b.", x = kreise_type) == FALSE) %>% 
  
  # Pad district code with leading 0s so that all are 5 characters long:
  mutate(kreise_code = stringr::str_pad(string = kreise_code, 
                                        width = 5, 
                                        side = "left", 
                                        pad = "0")) %>% 
  
  # Select only code and population:
  dplyr::select(kreise_code, 
                kreise_name, 
                kreise_type, 
                kreise_pop)

###############################################################
# Import population data for Berlin sub-districts:

# Import bundesland two-letter codes and population for Berlin sub-districts:
pop_berlin <- rio::import(here("resources", 
                               "SBD_pop_01_12_BERLIN_flat.csv"))

# Fix encoding of county names:
Encoding(pop_berlin$`Stadt- o. Landkreis`) <- "latin1"

# Complete preparation of data:
pop_berlin <- pop_berlin %>% 

  # Clean column names:
  janitor::clean_names() %>% 
  
  # Select just the columns for the lookup:
  dplyr::select(bundesland, 
                stadt_o_landkreis, 
                kreiskennziffer, 
                population) %>% 
  
  # Rename to match final county column in line list:
  rename(bundesland_lettercode = bundesland, 
         area_name = stadt_o_landkreis, 
         area_code = kreiskennziffer,
         area_pop = population) %>% 
  
  # Pad area codes so that all are 5 characters long:
  mutate(area_code = stringr::str_pad(string = area_code,
                                      width = 5, 
                                      side = "left", 
                                      pad = "0"))

```



```{r import_shapefiles}

#####################################################
# Import shape files for bundesland (federal states):

# Suppress output from R markdown:
invisible(
  capture.output(
    
    # Import shapefiles for bundesland (federal states):
    sf_bundesland <- sf::st_read(dsn = here::here("resources", 
                                            "shapefiles", 
                                            "vg2500_bld.shp"))
  )
)

# Correct formatting of German accents:
Encoding(sf_bundesland$GEN) <- "latin1" 
 
# Complete shapefile prep:
sf_bundesland <- sf_bundesland %>% 
  
  # Clean column names:
  janitor::clean_names() %>% 
   
  # Rename key variables:
  rename(bundesland_numcode = "rs", 
         bundesland_name = "gen") %>% 
  
  # Add population data:
  left_join(pop_bundesland %>% select(-bundesland_name)) %>% 
  
  # Remove unnecessary columns:
  select(-use, -rs_alt) %>% 
  
  # Relocate lettercode:
  relocate(bundesland_lettercode, .after = bundesland_numcode)


###########################################################
# Import shape files for kreise (administrative districts):

# Suppress output from R markdown:
invisible(
  capture.output(
    
    # Import Kreise shape files:
    sf_kreise <- sf::st_read(dsn = here::here("resources",
                                        "shapefiles", 
                                        "vg2500_krs.shp"))
  )
)

# Correct formatting of German accents:
Encoding(sf_kreise$GEN) <- "latin1" 
 
# Complete shapefile prep:
sf_kreise <- sf_kreise %>% 
  
  # Clean column names:
  janitor::clean_names() %>% 
  
  # Rename key variables:
  rename(kreise_code = "rs",
         kreise_name = "gen") %>%
  
  # Recode incorrect code for Gottingen (03152 -> 03159):
  mutate(kreise_code = ifelse(test = kreise_code == "03152",
                              yes = "03159", 
                              no = kreise_code)) %>% 
  
  # Add population data:
  left_join(pop_kreise %>% select(kreise_code, kreise_pop)) %>% 
  
  # Add missing population for Eisenach (16056):
  mutate(kreise_pop = ifelse(kreise_code == "16056", 
                             pop_berlin %>% 
                               filter(area_code == "16056") %>%
                               pull(area_pop), 
                             kreise_pop)) %>% 
  
  # Remove unnecessary columns:
  select(-use, -rs_alt)

# Population missing for Kreise 'Osterode am Harz' code 03156

################################################################################
# Import shape files for Berlin sub-districts:

# Suppress output from R markdown:
invisible(
  capture.output(
    
    # Import Berlin sub-district shape files:
    sf_berlin <- sf::st_read(dsn = here::here("resources",
                                        "shapefiles", 
                                        "SKLKBerlinBez.shp")) %>% 
      
      # Rename key variables:
      rename(berlin_district_code = "LKID",
             berlin_district_name = "byLKmeldel") %>%
      
      # Filter for only Berlin sub-districts:
      filter(grepl(pattern = "Berlin", x = berlin_district_name)) %>% 
      
      # Select only code and name columns:
      select(berlin_district_code, berlin_district_name) %>% 
      
      # Remove prefix from Berlin district names:
      mutate(berlin_district_name = gsub(pattern = "^SK Berlin ", 
                                         replacement = "", 
                                         x = berlin_district_name)) %>% 
      
      # Change Berlin district_code to character:
      mutate(berlin_district_code = as.character(berlin_district_code)) %>% 
      
      # Add population data:
      left_join(pop_berlin %>% select(area_code, area_pop), 
                by = join_by(berlin_district_code == area_code)) %>% 
      
      # Rename area pop:
      rename(berlin_district_pop = area_pop) %>% 
      
      # Transform to WGS 84 EPSG 4326 (GIS longitude / latitude):
      # Coordinate reference system must match the other two shapefiles
      st_transform(crs = st_crs(sf_kreise))

  )
)

```


```{r import_kauffmannwhite}

# Import the Kauffmann-White Salmonella serotype lookup table:
kw_lookup <- rio::import(here("resources", 
                              "Kauffmann_White_classification.xlsx")) %>% 
  
  clean_names()

```



```{r import_outbreakname, eval = params$outbreak_declared}

# Import the name of the cluster from the line list:
outbreakname <- rio::import(tmp_current, 
                           sheet = "Linelist", 
                           range = "B1", 
                           col_names = FALSE) %>% 
  
  # Pull the value out of the data.frame:
  pull()


# Update the title of this report:

### Note: updated yaml header is put in the main body (outside chunk) below.

```


```{r import_reportdate}

# Import the date of latest updates from the line list:
latest_update <- rio::import(tmp_current, 
                             sheet = "Linelist", 
                             range = "B3", 
                             col_names = FALSE) %>% 
  
  # Pull the value out of the data.frame:
  pull()


# Update the date of this report:

### Note: updated yaml header is put in the main body (outside chunk) below.

```


---
date: "`r format(latest_update, format = '<br/>Berichtsstand %d.%m.%Y')`"
---

```{r import_outbreakdate, eval = params$outbreak_declared}

############################################################################
# GET OUTBREAK DECLARATION DATE FROM LINELIST SHEET

# Import the outbreak declaration date:
outbreak_decdate <- rio::import(tmp_current, 
                                sheet = "Linelist", 
                                range = "J1", 
                                col_names = FALSE) %>% 
  
  # Extract the date to a named object:
  pull()


```


```{r import_clusterdate}

############################################################################
# GET CLUSTER NOTIFICATION DATE FROM LINELIST SHEET


# Import the cluster notification date:
cluster_decdate <- rio::import(tmp_current, 
                               sheet = "Linelist", 
                               range = "J2", 
                               col_names = FALSE) %>% 
  
  # Extract the date to a named object:
  pull()


```


```{r import_data_dictionary}

#####################################################################
# CREATE LOOKUP TABLE FOR SHORT VARIABLE NAMES

# Import data dictionary / lookup table for shorter variable names:
dictionary <- rio::import(here("resources", "Datenlexikon.csv")) 

```


```{r import_linelist_current}

#####################################################################
# IMPORT AND CLEAN CURRENT LINELIST

# Import raw linelist:
linelist <- rio::import(tmp_current, 
                        sheet = "Linelist", 
                        skip = 5, 
                        .name_repair = "minimal") %>% 
  
  # Clean the linelist with the function:
  clean_data()


# Note:
# - In the Shiny app, varname_code will be used to assign short names to the
# variables required to make this report only (for testing all imported 
# variables are assigned a short name, including hidden variables).
# - The short variable names are used in the rest of this r markdown
# - If you want to check what the short names correspond to, look them up in 
# the Datenlexikon.csv file which is in the resources folder.
# - If you add any new variables to the linelist that you want to use in the 
# report, make sure the cleaned names are first added to Datenlexikon.csv (the 
# order doesn't matter, they can be added at the end). 


```


```{r import_linelist_previous, eval = params$previous_report}

#####################################################################
# OPTIONALLY IMPORT AND CLEAN PREVIOUS LINELIST

# Imported only if params$previous_report == TRUE

# Import line list for previous sitrep:
linelist_old <- rio::import(tmp_previous, 
                        sheet = "Linelist", 
                        skip = 5, 
                        .name_repair = "minimal") %>% 
  
  # Clean the line list with the function:
  clean_data()

```



```{r get_outbreak_details}

############################################################################
# In this section we extract all the info needed for the front page summary
############################################################################

# The summary text is divided into the following sections:
############################################################################
### 01. OVERVIEW:

### --- a. cluster_decdate 
### --- b. clustername
### --- c. outbreak_decdate
### --- d. outbreakname

# a. Format cluster_decdate (WGS cluster NRZ notification date):
cluster_decdate <- format(cluster_decdate, format = "%d. %B %Y")


# b. Extract clustername from linelist:
clustername <- linelist %>% 
  
  filter(!is.na(isolate_ngsclusterid)) %>% 
  
  distinct(isolate_ngsclusterid) %>% 
  
  stringr::str_flatten_comma(last = " und ", na.rm = TRUE)


# c. Format outbreak_decdate if an outbreak has been declared:
if(params$outbreak_declared){
  
  outbreak_decdate <- format(outbreak_decdate, format = "%d. %B %Y")
  
}


# d. outbreakname:
# Already defined above


############################################################################
### 02. VALUES THAT REQUIRE CONDITIONAL FORMATTING:

# Use execsum function to calculate values that require conditional formatting:
currentvals <- execsum(data = linelist, suffix = "current")

# If a previous line list was uploaded, run the execsum function on it too:
if(params$previous_report == TRUE){
  
  previousvals <- execsum(data = linelist_old, suffix = "previous")
  
} else {
  
  previousvals <- execsum(data = linelist, suffix = "previous")
  
}


############################################################################
### 03. CREATE COMPARISON TABLE FOR FORMATTING:

# Compare bundesland in current and previous line lists:
if(params$previous_report == TRUE){
  
  bundeslist <- compare_bundesland(current = linelist, 
                                   previous = linelist_old)
  
} else {
  
  bundeslist <- compare_bundesland(current = linelist, 
                                   previous = linelist)
  
}


# Create the compare table:
compare <- compare_table(current = currentvals,
                         previous = previousvals, 
                         bundeslist = bundeslist) %>% 
  
  # Ensure latest update date always gets formatted:
  mutate(change = if_else(name == "latest_update", 
                          TRUE, 
                          change))


############################################################################
### 04. EXTRACT NUMBERS FOR EACH CASE DEFINITION:

# Get n_confirmed from compare table:
n_confirmed <- compare %>% 
  filter(name == "n_confirmed") %>% 
  select(value_current) %>% 
  pull() %>% 
  as.numeric()

# Create condition to show or hide confirmed cases statement:
if(n_confirmed > 0){showconf <- TRUE} else {showconf <- FALSE}

# Get n_probable from compare table:
n_probable <- compare %>% 
  filter(name == "n_probable") %>% 
  select(value_current) %>% 
  pull() %>% 
  as.numeric()

# Create condition to show or hide confirmed cases statement:
if(n_probable > 0){showprob <- TRUE} else {showprob <- FALSE}

# Get n_possible from compare table:
n_possible <- compare %>% 
  filter(name == "n_possible") %>% 
  select(value_current) %>% 
  pull() %>% 
  as.numeric()

# Create condition to show or hide possible cases statement:
if(n_possible > 0){showposs <- TRUE} else {showposs <- FALSE}

# Calculate number of isolates for which sequencing is pending:
n_seqpending <- linelist %>% 
  
  filter(isolate_ngsstatus == "vorgemerkt") %>% 
  
  count() %>% 
  
  pull()

# Create condition to show or hide sequence pending statement:
if(n_seqpending > 0){showseqpending <- TRUE} else {showseqpending <- FALSE}

# Create condition to show or hide new bundesland statement:
if(bundeslist$newstate_n > 0){
  showbundnew <- TRUE
} else {showbundnew <- FALSE}

# Create condition to show or hide bundesland with new cases statement:
if(bundeslist$newcases_n > 0){
  showbundrise <- TRUE
} else {showbundrise <- FALSE}

# Create condition to show or hide bundesland with no change statement:
if(bundeslist$nochange_n > 0){
  showbundsame <- TRUE
} else {showbundsame <- FALSE}

# Extract number of salmonellosis deaths:
n_deaths_salm <- compare %>% 
  filter(name == "n_deaths_salmonella") %>% 
  select(value_current) %>% 
  pull() %>% 
  as.numeric()

# Extract number of other deaths:
n_deaths_other <- compare %>% 
  filter(name == "n_deaths_other") %>% 
  select(value_current) %>% 
  pull() %>% 
  as.numeric()

# Create condition to show 'no deaths' statement or not:
if(n_deaths_salm + n_deaths_other == 0){
  shownodeaths <- TRUE
  nodeath_statement <- "Bislang wurden dem RKI keine Todesfälle übermittelt."
} else {
  shownodeaths <- FALSE
}

# Create condition for singular salmonella deaths:
if(n_deaths_salm == 1){
  nds_caseword <- "Todesfall"
} else {
  nds_caseword <- "Todesfälle"
}

# Create condition for singular other deaths:
if(n_deaths_other == 1){
  ndo_caseword <- "Todesfall"
} else {
  ndo_caseword <- "Todesfälle"
}

############################################################################
### 05. DEFINE REPORT TITLE CONDITIONALLY ON CLUSTER OR OUTBREAK:

# Update report title to outbreak or cluster as defined by outbreak_declared:
report_title <- ifelse(
  params$outbreak_declared == TRUE, 
  paste0('Zusammenfassender Bericht über  Salmonellenausbrüche:\n', 
         outbreakname),  
  paste0('Zusammenfassender Bericht über Salmonella NGS-Cluster:\n', 
         clustername))

```


---
title: "`r report_title`"
---

`r ftext("Zusammenfassung", prop = nontoc_header)`
<br>


:::{custom-style="Colour_background"}

<br>

`r ftext("Überblick", prop = abstract_header)`

Am `r cluster_decdate` hat das Nationale Referenzzentrum für Salmonellen und andere bakterielle Enteritiserreger ein neues NGS-Cluster mit der Bezeichnung `r clustername` identifiziert. Das Cluster gehört `r colourtext("serovars")` (`r colourtext("serogroups")`) an.  `r if(params$outbreak_declared){'Aufgrund der zeitlichen und geografischen Häufung wurde dieses Geschehen vom RKI am '}``r if(params$outbreak_declared){outbreak_decdate}` `r if(params$outbreak_declared){' als Ausbruch mit der Bezeichnung '}``r if(params$outbreak_declared){outbreakname}` `r if(params$outbreak_declared){' deklariert.'}`


Bis zum `r colourtext('latest_update')` wurden insgesamt `r colourtext('n_cases')` Ausbruchsfälle diesem Geschehen zugeordnet. 


<!-- `r if(showconf){"Davon sind "}``r if(showconf){colourtext('n_confirmed')}``r if(showconf){" als bestätigte Fälle kategorisiert."}` `r if(showprob){colourtext('n_probable')}``r if(showprob){" sind als wahrscheinlich kategorisiert."}` `r if(showposs){colourtext('n_possible')}``r if(showposs){" sind aufgrund der Serotypisierungsergebnisse als mögliche Fälle kategorisiert."}` `r if(showseqpending){'Für '}``r if(showseqpending){colourtext('n_seqpending')}``r if(showseqpending){' Ausbruchsfall / Ausbruchsfälle stehen die Sequenzierungsergebnisse noch aus.'}` -->


Die Ausbruchsfälle stammen aus `r colourtext("n_bundesland")` verschiedenen Bundesländern. Dazu gehören `r if(showbundnew){colourtext('bundesland_new_n')}``r if(showbundnew){" Bundesländer, die zum ersten Mal Fälle melden ("}``r if(showbundnew){colourtext('bundesland_new_statement')}``r if(showbundnew){")"}` `r if(showbundrise){colourtext('bundesland_increased_n')}``r if(showbundrise){" mit neuen Fällen ("}``r if(showbundrise){colourtext('bundesland_increased_statement')}``r if(showbundrise){")"}` `r if((showbundnew | showbundrise) & showbundsame){" und "}``r if(showbundsame & params$previous_report){colourtext('bundesland_nochange_n')}``r if(showbundsame & params$previous_report){" ohne Veränderung der Fallzahlen seit dem letzten Bericht: "}``r if(showbundsame){colourtext('bundesland_nochange_statement')}`.


Die Fälle sind über einen Zeitraum von `r colourtext("n_totalweeks")` Kalenderwochen aufgetreten. Der letzte Fall wurde am `r colourtext("max_reportdate")` gemeldet und der letzte übermittelte Erkrankungsbeginn ist der `r colourtext("max_onsetdate")`. 


`r colourtext("statement_sexratio")` (`r colourtext("n_female")` x weiblich *vs.* `r colourtext("n_male")` x männlich). `r colourtext("statement_adultchild")` (mittleres Alter: `r colourtext("median_age")` Jahre, Spanne: `r colourtext("range_age")`, IQR: `r colourtext("iqr_age")`). Für `r colourtext("n_hosp_info")` Ausbruchsfälle liegen ausreichende Informationen zu Krankenhausaufenthalten vor. Davon wurden `r colourtext("n_hosp_salmonella")` aufgrund der Salmonellose und `r colourtext("n_hosp_other")` aufgrund anderer Ursachen ins Krankenhaus eingeliefert. `r if(shownodeaths){nodeath_statement}``r if(!shownodeaths){"Es gab "}``r if(!shownodeaths){colourtext('n_deaths_salmonella')}` `r if(!shownodeaths){nds_caseword}``r if(!shownodeaths){" aufgrund der Salmonellose "}``r if(!shownodeaths & n_deaths_salm > 0){"("}``r if(!shownodeaths & n_deaths_salm > 0){colourtext('deathsalm_different')}``r if(!shownodeaths & n_deaths_salm > 0){colourtext('deathsalm_same')}``r if(!shownodeaths & n_deaths_salm > 0){")"}` `r if(!shownodeaths){" und "}``r if(!shownodeaths){colourtext('n_deaths_other')}``r if(!shownodeaths){" Todesfall aufgrund anderer Ursachen"}``r if(!shownodeaths & n_deaths_other > 0){" ("}``r if(!shownodeaths & n_deaths_other > 0){colourtext('deathother_different')}``r if(!shownodeaths & n_deaths_other > 0){colourtext('deathother_same')}``r if(!shownodeaths & n_deaths_other > 0){")"}``r if(!shownodeaths){"."}` 


`r ftext("Bewertung der Lage", prop = abstract_header)`

`r ftext("[Fügen Sie hier zusätzlichen Text ein.]", prop = placeholder)`

<br>

:::

<br>
<br>


```{r outbreak_infotable}

# Extract total number of cases:
n_cases <- compare %>% 
  
  filter(name == "n_cases") %>% 
  
  pull(value_current)

# Extract serovar:
serovars = linelist %>% 
  
  distinct(serovar_final) %>% 
  
  stringr::str_flatten_comma(last = " und ", na.rm = TRUE)


# Create small table to hold the information:
if(params$outbreak_declared){
  
  infotab <- data.frame(
  
  # Create labels according to report template:
  infolabel = c("Bezeichnung des Ausbruches im RKI:", 
                "Enthält folgende NGS-Cluster:", 
                "In diesem Ausbruch enthaltene Serovare:",
                "Anzahl der Ausbruchsfälle in Deutschland:"), 
  
  # Supply already calculated values:
  infovalue = c(outbreakname, 
                clustername,
                serovars,
                as.character(n_cases))
  )

} else {
  
  infotab <- data.frame(
  
  # Create labels according to report template:
  infolabel = c("Enthält folgende NGS-Cluster:", 
                "In diesem Ausbruch enthaltene Serovare:",
                "Anzahl der Ausbruchsfälle in Deutschland:"), 
  
  # Supply already calculated values:
  infovalue = c(clustername,
                serovars,
                as.character(n_cases))
  )

}


  
# Process as flextable:  
infotab <- infotab %>% 
  
  # Convert to a flextable:
  flextable() %>% 
  
  # Convert to box theme:
  theme_box() %>% 
  
  # Change colour of all rows except last one to pale blue:
  bg(i = 1:(nrow(infotab) - 1), 
     part = "body", 
     bg = "#C6D9F1") %>% 
  
  # Change colour of last row to dark blue:
  bg(i = ~ infovalue == n_cases, 
     part = "body",
     bg = "#4F81BD") %>% 
  
  # Change font colour of header to white:
  color(i = ~ infovalue == n_cases,
        part = "body",
        color = "white") %>% 
  
  # Right align values column:
  align(j = ~ infovalue,
        align = "right", 
        part = "body") %>% 
  
  # Make values bold:
  bold(j = ~ infovalue, 
       part = "body") %>% 
  
  # Change font size to match front page:
  fontsize(size = 10, part = "body") %>% 
  
  # Remove borders:
  border_remove() %>% 
  
  # Delete header row:
  delete_part(part = "header") %>% 
  
  # Set table width:
  width(., width = pagewidth_cm/ncol(infotab), unit = "cm")


### PRINT THE OUTBREAK INFO TABLE:
infotab

```


\newpage


`r ftext("Inhaltsverzeichnis", prop = nontoc_header)`

```{r create_toc}

# Insert table of contents:
officer::block_toc()

```


\newpage


# Falldefinitionen

```{r import_casedefinition}

######################################################
# Import case definition date: 
casedef_date <- rio::import(tmp_current, 
                             sheet = "Falldefinition", 
                             range = "B11", 
                             col_names = FALSE) %>% 
  
  # Pull the value out of the data.frame:
  pull()


######################################################
# Import the case definition from the line list:
casedef <- rio::import(tmp_current, 
                       sheet = "Falldefinition", 
                       skip = 12, 
                       col_names = FALSE) %>% 
  
  # Clean column names:
  janitor::clean_names() %>% 
  
  # Remove rows with NA (oder):
  drop_na(x1) %>% 
  
  # Reformat exclusions (Ausschluss) so that case type is column x5:
  mutate(x5 = if_else(condition = is.na(x5), 
                      true = "Ausschluss", 
                      false = x5)) %>% 
  
  # Remove the Ausschluss title row:
  filter(! x1 %in% "Ausschluss") %>% 
  
  # Remove numbers for exclusions:
  mutate(across(.cols = c(x2,x3,x4), 
                .fns = function(x) str_replace_all(x, "^[:digit:]+$", ""))) %>% 
  
  # Create combined cols for each definition:
  unite(., col = "Definition",  x1, x2, x3, sep = " ", na.rm = TRUE) %>% 
  
  # Rename the column with case type:
  rename(Falltyp = x5) %>% 
  
  # Select columns and reorder:
  dplyr::select(Falltyp, Definition) %>% 
  
  # Change falltyp to title case:
  mutate(Falltyp = tools::toTitleCase(tolower(Falltyp))) %>% 
  
  # Change 'outbreak' to 'case' in case type column:
  mutate(Falltyp = gsub(pattern = "Ausbruchsfall", 
                        replacement = "Fall", 
                        x = Falltyp)) %>% 
  
  # Convert to flextable:
  qflextable() %>% 
  
  # Change theme:
  theme_box() %>% 
  
  # Merge repeat rows for Falltyp:
  merge_v(j = ~Falltyp) %>% 
  
  # Make Falltyp values bold:
  bold(j = ~Falltyp) %>% 
  
  # Add pale green for confirmed case definitions:
  bg(i = ~ Falltyp == "Bestätigter Fall", 
     bg = "palegreen", 
     part = "body") %>% 
  
  # Add orange for probable case definitions:
  bg(i = ~ Falltyp == "Wahrscheinlicher Fall", 
     bg = "orange", 
     part = "body") %>% 
  
  # Add pink for possible case definitions:
  bg(i = ~ Falltyp == "Möglicher Fall", 
     bg = "lightpink", 
     part = "body") %>% 
  
  # Add grey for exclusion criteria:
  bg(i = ~ Falltyp == "Ausschluss", 
     bg = "lightgrey", 
     part = "body") %>% 
  
  # Change background colour of header row:
  bg(bg = "#4F81BD", 
     part = "header") %>% 
  
  # Change font colour of header to white:
  color(color = "white", 
        part = "header") %>% 
  
  # Fit columns to text:
  set_table_properties(layout = "autofit")


### PRINT THE CASE DEFINITIONS TABLE:
casedef


### Note: Use set_table_properties(layout = "autofit") and not autofit() when
# knitting to an MS word document.  This will fit the table to the margins.
# autofit() will result in wide columns extending past the document page limit.

```


\pagebreak

# Deskriptive Analyse des Ausbruchs


## Geografische Verteilung der Fälle


**Betroffene Bundesländer** (*n* = `r colourtext("n_bundesland")`):  

```{r fedstates_list}

# Extract the list of federal states with case numbers from compare table:
fedstates_list <- compare %>% 
  
  filter(name == "statement_bundesland") %>% 
  
  pull(value_current)

```


`r ftext(fedstates_list, prop = nstyle)`


```{r geodist_table, tab.cap = "Fallzahlen nach Bundesland des Wohnsitzes und/oder der Exposition"}

# Get full list of two-letter codes for federal state:
fedstate_codes <- unique(pop_berlin$bundesland_lettercode)

# Group by Bundesland:
fedstate_restab <- linelist %>%  

  group_by(bundesland_lettercode) %>% 
  
  summarise(n_res = n())

# Group by exposelocation_final:
fedstate_exptab <- linelist %>% 
  
  group_by(exposelocation_final) %>% 
  
  summarise(n_exp = n())

# Merge the two tables:
fedstatetab <- full_join(
  x = fedstate_restab,
  y = fedstate_exptab,
  by = join_by(bundesland_lettercode == exposelocation_final)) %>% 
  
  # Replace NAs in numeric columns with 0:
  mutate(across(where(is.numeric), ~replace_na(.x, 0))) %>% 
  
  # Rename bundesland:
  rename(bundesland = bundesland_lettercode) %>% 
  
  # Replace NA with unbekannt for bundesland column:
  mutate(bundesland = replace_na(bundesland, "unbekannt")) %>% 
  
  # Pivot longer to start transposition of table:
  pivot_longer(-bundesland) %>%
  
  # Complete the transposition:
  pivot_wider(name, names_from = "bundesland", values_from = "value") %>% 
  
  # Add the sum column:
  janitor::adorn_totals(where = "col", name = "Summe") %>% 
  
  # Rename row headers:
  mutate(name = case_when(
    name == "n_res" ~ "Wohnort",
    name == "n_exp" ~ "Exposition"
  ))

  
# Check which federal states are not yet included in the table:
fs2add <- fedstate_codes[!fedstate_codes %in% names(fedstatetab)]


# Add new columns dynamically to table:
for(x in fs2add) fedstatetab <- fedstatetab %>% 
  
  # Give them all 0 as values:
  mutate(!! x := 0)


# Now re-order columns: 
fedstatetab <- fedstatetab %>% 
  
  relocate(any_of(c("name", fedstate_codes, "unbekannt", "Ausland", "Summe")))


# Convert to flextable and format:
fedstateprint <- fedstatetab %>% 

  # Convert to flextable:
  qflextable() %>% 
  
  # Allow markdown formatting to hyperlink email addresses:
  ftExtra::colformat_md(.sep = "\n") %>% 
  
  # Rename row header:
  set_header_labels(name = "Fallzahlen pro") %>% 
  
  # Remove all borders with a theme:
  theme_box() %>% 
  
  # Make column titles bold:
  bold(part = "header") %>% 
  
  # Make last column italic:
  italic(j = ncol(fedstatetab), 
         part = "all") %>% 
  
  # Make first and last column body text bold:
  bold(j = c(1, ncol(fedstatetab)), 
       part = "body") %>% 
  
  # Make all the middle columns white text:
  color(j = 2:(ncol(fedstatetab) - 1), 
        color = "white", 
        part = "header") %>% 
  
  # Change background colour of header row:
  bg(bg = "#4F81BD", 
     part = "header") %>%
  
  # Change the font size for whole table to 10 pt:
  fontsize(size = 10, part = "all") %>% 

  # Rotate headers of last three columns:
  rotate(j = (ncol(fedstatetab)-2):ncol(fedstatetab),
         align = "center", 
         rotation = "btlr",
         part = "header") %>% 
  
  # Align the rest of the column headers to bottom:
  valign(j = 1:(ncol(fedstatetab)-3), 
         valign = "center", 
         part = "header") %>% 
  
  # Adjust height of header row:
  height(height = 1.0, part = "header") %>% 
  
  # Adjust rule:
  hrule(i = 1, rule = "exact", part = "header") %>% 
  
  # Fit columns to text:
  set_table_properties(layout = "autofit") 


# Print the table:
fedstateprint

```

<br>


```{r bundesland_bargraph, fig.cap = "Vergleich der Inzidenzen (Fallzahlen/100.000 Einwohner) pro Bundesland"}

# Create table of cases by federal state:
fedstate_incidence <- linelist %>% 
  
  # Remove NAs in bundesland:
  drop_na(bundesland_lettercode) %>% 
  
  # Group by bundesland:
  group_by(bundesland_lettercode) %>% 
  
  # Select bundesland:
  select(bundesland_lettercode) %>%
  
  # Rename bundesland:
  rename(bundesland = bundesland_lettercode) %>% 
  
  # Summarise number of cases by bundesland:
  summarise(n_cases = n()) %>% 
  
  # Add rows for missing bundeslands (assign 0 cases):
  tibble::add_row(bundesland = fs2add, n_cases = 0) %>% 
  
  # Join population to table:
  left_join(pop_bundesland %>% select(hexcode, 
                                      bundesland_lettercode, 
                                      bundesland_pop), 
            join_by(bundesland == bundesland_lettercode)) %>% 
  
  # Put hexcode at beginning:
  relocate(hexcode, .after = bundesland) %>% 
  
  # Create bundesland incidence per 100,000 column:
  mutate(incidence = round((n_cases / (bundesland_pop/100000)), digits = 2)) %>% 
  
  # Arrange table by decreasing incidence, then bundesland (alphabetic):
  arrange(desc(incidence), bundesland) %>% 
  
  {
    # Set plotting aesthetics (x and y axes):
    ggplot(., mapping = aes(x = reorder(bundesland, incidence),
                            y = incidence, 
                            fill = hexcode)) +
      
      # Use stat = identity to use calculated incidence as y axis:
      geom_bar(stat = "identity") + 
      
      # Use colour values assigned as fill in aes for fill colour:
      scale_fill_identity() +
  
      # Set y axis breaks and limits dynamically:
      scale_y_continuous(
        limits = c(0.00, max(.$incidence) + 0.01), 
        breaks = seq(from = 0.00, 
                     to = max(.$incidence),
                     by = 0.02), 
        expand = c(0,0)) + # Get rid of white space below min limit on x axis
      
      # Update x axis label:
      xlab("Bundesland des Wohnortes\n") +
  
      # Update y axis label:
      ylab("\nInzidenz (Ausbruchsfälle/100.000 Einwohner)") +

      # Flip graph on its side:
      coord_flip() +
      
      # Apply plain theme:
      theme_classic() +
      
      # Make axis titles bold:
      theme(axis.title = element_text(face = "bold"))
    
  }

############################
# Print bar graph:
fedstate_incidence

```



\pagebreak


```{r map_data_prep}

##########################################################
# DATA PREP FOR INCIDENCE CALCULATIONS

# Create table of cases by kreise:
kreise_tab <- linelist %>%
  
  select(kreise_final, kreise_code) %>% 
  
  drop_na(kreise_final) %>% 
  
  group_by(kreise_final, kreise_code) %>% 
  
  summarise(n_cases = n())


# Calculate number of sub-districts in Berlin with cases:
berlin_districts <- linelist %>% 
  
  filter(!is.na(berlin_district_name)) %>%
  
  distinct(berlin_district_name) %>% 
  
  count() %>% 
  
  pull()

# Create conditional statement for Berlin districts:
if(berlin_districts > 1){
  showberlin <- TRUE
} else {
  showberlin <- FALSE
}

# Conditionally create Berlin sub-districts table:
if(berlin_districts > 1){
  
  # Create table of cases by Berlin sub-district:
  berlin_tab <- linelist %>% 
    
    select(berlin_district_name, berlin_district_code) %>% 
    
    drop_na(berlin_district_name) %>% 
    
    group_by(berlin_district_name, berlin_district_code) %>% 
    
    summarise(n_cases = n())
}


##########################################################
# INCIDENCE CALCULATIONS:

##### Create incidence table:
kreise_incidence <- sf_kreise %>% 
  
  # Join county table for case numbers:
  left_join(kreise_tab %>% select(-kreise_final)) %>% 
  
  # Remove areas with no population figures:
  drop_na(kreise_pop) %>% 
  
  # Convert county population from character to numeric:
  mutate(kreise_pop = as.numeric(kreise_pop)) %>% 
  
  # Calculate incidence per 100000 population:
  mutate(incidence = round((n_cases / (kreise_pop/100000)), digits = 3)) %>% 
  
  # Put incidence after n_cases to make it easier to review:
  relocate(incidence, .after = n_cases) %>% 
  
  # Arrange by incidence in descending order:
  arrange(desc(incidence)) %>% 
  
  # Create column with incidence break labels:
  mutate(incidence_breaks = ibreak_labeller(x = incidence)) 


##########################################################
# DEFINE COLOUR PALATTE:

# Define colour scheme:
mypal <- grDevices::hcl.colors(n = 5, 
                               palette = "Blues", 
                               rev = TRUE)[2:5]

# Update mypal to include white for NAs:
mypal <- c("white", mypal)

# Extract the incidence break labels:
ibreak_labels <- kreise_incidence %>% 
  
  # Remove geometry:
  st_drop_geometry() %>% 
  
  # Remove NAs from incidence breaks:
  drop_na(incidence_breaks) %>% 
  
  # Fetch distinct values for incidence breaks:
  distinct(incidence_breaks) %>% 
  
  # Pull the distinct values:
  pull() %>% 
  
  # Reverse the order (from descending to ascending):
  rev()


# Create the colour palette:
map_pal <- mypal[1:length(ibreak_labels)]


# Put the colours and labels together:
mapcolours <- setNames(object = map_pal, 
                       nm = ibreak_labels)

##########################################################
# CLASS INTERVAL METHOD LABEL:

# Improve description of class interval method from params:
if(params$map_breaks == "fisher"){
  mapinterval_label = "Fisher Jenks"
} else {
  mapinterval_label = params$map_breaks
}

```


```{r kreise_map, fig.cap = "Stadt- oder Landkreise mit Ausbruchsfällen", fig.height = 6}

##########################################################
# CHOROPLETH MAP OF INCIDENCE BY KREISE:

#Create the map:
kreise_map <- kreise_incidence %>% 
  
  # Start the plot:
  ggplot2::ggplot() + 
  
  # Add the sf_kreise geom with incidence breaks:
  geom_sf(
    mapping = aes(geometry = geometry, fill = incidence_breaks), 
    colour = "darkgrey", 
    size = 0.5, 
    na.rm = TRUE) +
  
  # Add the sf_bundesland geom for bundesland borders:
  geom_sf(data = sf_bundesland,
          fill = alpha("white", 0), 
          colour = "black", 
          size = 1) +

  # Specify how the fill colours should be set:
  scale_fill_manual(
    values = mapcolours,
    na.translate = FALSE
    ) +

  # Set the legend title and colour order in the legend:
  guides(fill = guide_legend(
    title = expression(atop(bold("Inzidenz"), "(Fallzahlen/100.000 Einwohner)")), 
    reverse = FALSE)) + 
  
  # Apply the ggplot theme_nothing to remove grid lines:
  ggmap::theme_nothing(legend = TRUE) +
  
  # Add footnote with class interval method used:
  labs(caption = stringr::str_glue("Klassenintervallmethode: {mapinterval_label}")) +
  
  # Right-align caption to plot edge:
  theme(plot.caption.position = "plot")

#####################################
# Print county table and map:

kreise_map

```

<br>

```{r incidence_table, tab.cap = "Die Kreise mit der höchsten Inzidenz pro 100.000 Einwohner:"}

# Create table with top 5 Kreise according to incidence:
top5kreise <- kreise_incidence %>% 
  
  select(kreise_name, kreise_pop, n_cases, incidence) %>% 
  
  st_drop_geometry() %>% 
  
  slice_head(n = 5) %>% 
  
  qflextable() %>% 
  
  theme_booktabs() %>% 
  
  set_header_labels(
    kreise_name = "Kreis", 
    kreise_pop = "Bevölkerung",
    n_cases = "Anzahl der Fälle",
    incidence = "Inzidenz"
  ) %>% 
  
  bold(part = "header") %>% 
  
  # Fit columns to text:
  set_table_properties(layout = "autofit")

##########################################
# Print the table:

top5kreise

```



```{r berlin_map, eval = showberlin, fig.cap = "Berliner Bezirke mit Ausbruchsfällen (Fallzahlen/100.000 Einwohner)"}

# Add incidence to Berlin shapefile:
berlin_incidence <- sf_berlin %>% 
  
  # Join case numbers to shapefile:
  left_join(berlin_tab %>% select(-berlin_district_name)) %>% 
  
  # Make sure population is numeric:
  mutate(berlin_district_pop = as.numeric(berlin_district_pop)) %>% 
  
  # Calculate incidence per 100000 population:
  mutate(incidence = round((n_cases / (berlin_district_pop/100000)), 
                           digits = 3)) %>% 
  
  # Put incidence after n_cases to make it easier to review:
  relocate(incidence, .after = n_cases) %>% 
  
  # Arrange by incidence in descending order:
  arrange(desc(incidence)) 



# Create the map:
berlin_map <- berlin_incidence %>% 
  
  # Remove labels for areas with no incidence:
  mutate(berlin_district_name = case_when(
    is.na(incidence) ~ NA_character_,
    TRUE ~ berlin_district_name)) %>%

  # Start the plot:
  ggplot2::ggplot() + 
  
  # Add the sf_berlin geom:
  geom_sf(mapping = aes(fill = incidence), 
          colour = "black", 
          size = 1, 
          na.rm = TRUE) +
  
  # Specify how the fill colours should be set:
  scale_fill_distiller(palette = "Blues", 
                       na.value = "white", 
                       breaks = scales::breaks_pretty(),
                       labels = scales::label_number(decimal.mark = ","),
                       direction = 1) +

  # Add labels for Berlin districts with incidence > 0:
  ggsflabel::geom_sf_label_repel(mapping = aes(
    label = ifelse(!is.na(berlin_district_name), 
                   paste0(berlin_district_name, 
                          ": ", 
                          format(incidence, decimal.mark = ",")), 
                   NA_character_)),
    hjust = 0, 
    size = 3,
    box.padding = 0.35, 
    segment.size = 0.25, 
    nudge_x = 0.5,
    direction = 'y',
    na.rm = TRUE) +
  
  # Set the legend title and colour order in the legend:
  guides(fill = guide_legend(
    title = expression(atop(bold("Inzidenz"), "(pro 100.000 Einwohner)")), 
    reverse = FALSE)) + 
  
  # Apply the ggplot theme_nothing to remove grid lines:
  ggmap::theme_nothing(legend = TRUE)


####################
# Print map:

berlin_map

```


\pagebreak



## Zeitliche Verteilung der Fälle


```{r epicurve, fig.height = 5, fig.cap = "Epidemiekurve der Fälle: gesamter Ausbruch bis heute"}

# Define stratifier from params:
if(params$stratifier == "Bundesland"){
  
  # Define stratum as bundesland 2-letter code:
  strata <- "bundesland_lettercode"
  
  # Define name for legend:
  legendname <- "Bundesland"
  
  # Define colour palatte as bundesland hexcodes:
  colpalate <- setNames(object = c(pop_bundesland$hexcode, "lightgrey"), 
                        nm = c(pop_bundesland$bundesland_lettercode, "Unbekannt"))
  
} else if(params$stratifier == "Falldefinition"){
  
  # Define stratum as case definition:
  strata <- "case_definition"
  
  # Define name for legend:
  legendname <- "Falldefinition"
  
  # Define colour palatte from case definition table:
  colpalate <- setNames(object = c("palegreen", 
                                    "orange", 
                                    "lightpink",
                                   "lightgrey"), 
                         nm = c("Bestätigter", 
                                "Wahrscheinlicher", 
                                "Möglicher", 
                                "Unbekannt"))
}


######################################

### Define weekly breaks
week_breaks <- seq.Date(
  # Define minimum date in series (to start Monday before first case):
  from = floor_date(x = min(linelist$epicurve_dates, na.rm = TRUE), 
                    unit = "week", 
                    week_start = 1),
  # Define maximum date in series (to end Monday after latest case):
  to   = ceiling_date(x = max(linelist$epicurve_dates, na.rm = TRUE),
                      unit = "week", 
                      week_start = 1),
  # Define interval unit (aggregate in weeks):
  by   = "week")


### Define monthly breaks
month_breaks <- seq.Date(
  # Define minimum date in series (to start 1st day of month before 1st case):
  from = floor_date(x = min(linelist$epicurve_dates, na.rm = TRUE), 
                    unit = "month"),
  # Define maximum date in series (to end 1st day of month after last case):
  to   = ceiling_date(x = max(linelist$epicurve_dates, na.rm = TRUE),
                      unit = "month"),
  # Define interval unit (aggregate in months):
  by   = "month")


### Define quarter breaks
quarter_breaks <- seq.Date(
  # Define minimum date in series (to start 1st day of quarter before 1st case):
  from = floor_date(x = min(linelist$epicurve_dates, na.rm = TRUE), 
                    unit = "quarter"),
  # Define maximum date in series (to end 1st day of quarter after last case):
  to   = ceiling_date(x = max(linelist$epicurve_dates, na.rm = TRUE),
                      unit = "quarter"),
  # Define interval unit (aggregate in quarters):
  by   = "quarter")


######################################
# Conditional weeks or months:

# Extract total duration of outbreak in weeks:
n_totalweeks <- compare %>% 
  
  filter(name == "n_totalweeks") %>% 
  
  pull(value_current) %>% 
  
  as.numeric()

# Set main epicurve units to be year quarter, months or weeks:
if(n_totalweeks > 52){
  
  epicurve_interval <- "3 months"
  
  epicurve_order <- "epicurve_yearquarter"
  
  epicurve_breaks <- quarter_breaks
  
  epicurve_labels <- "%m"
  
} else if(n_totalweeks > 23 & n_totalweeks <= 52){
  
  epicurve_interval <- "month"
  
  epicurve_order <- "epicurve_yearmonth"
  
  epicurve_breaks <- month_breaks
  
  epicurve_labels <- "%b"
  
} else {
  
  epicurve_interval <- "week"
  
  epicurve_order <- "epicurve_yearweek"
  
  epicurve_breaks <- week_breaks
  
  epicurve_labels <- "%W"
  
}



######################################
# Create main epicurve:

# Prepare aggregate data:
epicurve <- linelist %>% 
  
  # Rename strata as strata:
  rename(strata = !!sym(strata)) %>% 
  
  # Add 'unknown' level to factor to represent NAs:
  mutate(strata = forcats::fct_na_value_to_level(
    f = strata,
    level = "Unbekannt")) %>%

  # Reorder levels of strata:
  mutate(strata = fct_relevel(.f = strata,
                              "Unbekannt",
                              after = Inf)) %>%
  
  # Arrange data by main epicurve_order and strata:
  arrange(!!sym(epicurve_order), strata) %>% 

  # Create grouping variable for epi squares:
  mutate(blocks = 1:n()) %>% 
  
  # Create plot:
  ggplot() +
  
  # Create geom:
  geom_histogram(mapping = aes(x = epicurve_dates, 
                               fill = strata, 
                               group = blocks), 
                 
                 just = 0.5,
                 breaks = epicurve_breaks,
                 closed = "left",
                 colour = "white") +
  
  # Set labels for date axis:
  scale_x_date(date_breaks = epicurve_interval,
               minor_breaks = NULL, 
               expand = expansion(add = 3),
               labels = label_date_axis) +
  
  # Remove space from bottom of x axis and present y axis labels as integers:
  scale_y_continuous(expand = c(0,0), 
                     breaks = ~round(unique(pretty(.)))) +             
  
  # Set colour palatte:
  scale_fill_manual(values = colpalate,
                    name = expr(bold(!!legendname)), 
                    na.value = "lightgrey") +
  
  # Add the axis and fill labels:
  labs(x = "\nDatum Erkrankungsbeginn (bzw. [Meldewoche o. Laborwoche - 1], wenn nicht bekannt)",
       y = "Anzahl der Fälle\n") +
  
  # aesthetic themes
  theme_classic() +
  
  # Make axis titles bold:
  theme(axis.title = element_text(face = "bold", size = 10), 
        axis.text.x = element_text(hjust = 0),
        axis.ticks.x = element_blank(), 
        legend.key.size = unit(0.5, 'cm'),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 9),
        legend.position = "top", 
        legend.justification = "left",
        legend.direction = "horizontal") +

  # Add markdown:
  theme(axis.text.x = ggtext::element_markdown()) +
  
  # Put legend in 9 columns:
  guides(fill = guide_legend(ncol = 10))



#####################
# Print the plot:

# Print main epicurve:
epicurve

# If outbreak > 23 weeks, print epicurve weeks subset:
if(n_totalweeks > 23){
  
  show_epicurve_detail = TRUE
  
} else {
  
  show_epicurve_detail = FALSE
  
}

```


```{r epicurve_detail, eval = show_epicurve_detail, fig.height = 5, fig.cap = "Epidemiekurve der Fälle: letzte 20 Wochen"}

################################################
# Add detailed epicurve if outbreak > 23 weeks:

###########################################
# Reset epicurve label prefix:
epicurve_labels <- "%W"

##########################################
# Calculate lower date limit for epicurve:

epicurve_minweek <- linelist %>% 
  
  select(epicurve_dates) %>% 
    
  drop_na() %>% 
    
  pull() %>% 
    
  max() - lubridate::weeks(20)
  
  
##################################
# Calculate subset of week breaks:

week_breaks_subset <- subset(week_breaks, 
                             week_breaks >= epicurve_minweek)
  
###############################
# Create weeks subset epicurve:

epicurve_detail <- linelist %>% 
  
  # Filter for epicurve dates greater than epicurve_minweek:
  filter(epicurve_dates >= epicurve_minweek) %>% 
    
  # Rename strata as strata:
  rename(strata = !!sym(strata)) %>% 
    
  # Add 'unknown' level to factor to represent NAs:
  mutate(strata = forcats::fct_na_value_to_level(
    f = strata,
    level = "Unbekannt")) %>%
    
  # Reorder levels of strata:
  mutate(strata = fct_relevel(.f = strata,
                              "Unbekannt",
                              after = Inf)) %>%
  
  # Arrange data by epicurve_weeks and strata:
  arrange(!!sym(epicurve_order), strata) %>% 
    
  # Create grouping variable for epi squares:
  mutate(blocks = 1:n()) %>% 
    
  # Create plot:
  ggplot() +
    
  # Create geom:
  geom_histogram(mapping = aes(x = epicurve_dates,
                               fill = strata, 
                               group = blocks), 
                 just = 0.5,
                 breaks = week_breaks_subset,
                 closed = "left",
                 colour = "white") +
    
  # Set labels for date axis:
  scale_x_date(date_breaks = "week",
               minor_breaks = NULL, 
               expand = expansion(add = 3),
               labels = label_date_axis) +
    
  # Remove space from bottom of x axis and present y axis labels as integers:
  scale_y_continuous(expand = c(0,0), 
                     breaks = ~round(unique(pretty(.)))) + 
  
  
    
  # Set colour palatte:
  scale_fill_manual(values = colpalate,
                    name = expr(bold(!!legendname)), 
                    na.value = "lightgrey") +
    
  # Add the axis and fill labels:
  labs(x = "\nDatum Erkrankungsbeginn (bzw. [Meldewoche o. Laborwoche - 1], wenn nicht bekannt)",
       y = "Anzahl der Fälle\n") +
  
  # aesthetic themes
  theme_classic() + 
    
  # Make axis titles bold:
  theme(axis.title = element_text(face = "bold", size = 10), 
        axis.text.x = element_text(hjust = 0),
        axis.ticks.x = element_blank(), 
        legend.key.size = unit(0.5, 'cm'),
        legend.title = element_text(size = 9),
        legend.text = element_text(size = 9),
        legend.position = "top", 
        legend.justification = "left",
        legend.direction = "horizontal") +
    
  # Add markdown:
  theme(axis.text.x = ggtext::element_markdown()) +
    
  # Put legend in 9 columns:
  guides(fill = guide_legend(ncol = 10))


# Print epicurve for last 20 weeks:
epicurve_detail

```



Der Median der Zahl der Fälle pro Woche (ohne Wochen ohne Fälle) beträgt `r compare$value_current[compare$name == "median_cpw"]`. Der früheste Fall trat am `r compare$value_current[compare$name == "min_epidate"]` auf (Kalenderwoche `r compare$value_current[compare$name == "min_epiweek"]`).  50 % der bisher gemeldeten Fälle (n = `r compare$value_current[compare$name == "n_50percentcases"]`) wurden bis zur Kalenderwoche `r compare$value_current[compare$name == "median_week"]` festgestellt, wobei die höchste wöchentliche Fallzahl (n = `r compare$value_current[compare$name == "max_cpw"]`) in der Kalenderwoche `r compare$value_current[compare$name == "max_cpw_week"]` auftrat. Der letzte Fall bzw. die letzten Fälle (n = `r compare$value_current[compare$name == "n_latestcases"]`) traten in der Kalenderwoche `r compare$value_current[compare$name == "max_epiweek"]` auf (letzte übermittelte Erkrankungsbeginn: `r compare$value_current[compare$name == "max_onsetdate"]`).


\pagebreak


## Beschreibung der Ausbruchsfälle nach Altergruppe und Geschlecht


```{r age_sex_pyramid, fig.cap = "Alter-Geschlecht-Pyramide"}

# Create named list for the colour palate:
sexcolours <- setNames(object = c("#B9CDE5",
                                  "#E6B9B8"), 
                        nm = c('männlich',
                               'weiblich'))

# Define what values of sex exist in the data:
sexindata <- linelist %>% 
  
  select(sex_final) %>% 
  
  pull() %>% 
  
  unique()

# Restrict sexcolours to what actually exists in the data:
sexcolours <- subset(sexcolours, names(sexcolours) %in% sexindata)


# Create the age sex pyramid:
agesexpyramid <- linelist %>% 
  
  drop_na(age_cat) %>% 
  
  drop_na(sex_final) %>% 
  
  apyramid::age_pyramid(age_group = "age_cat", 
                        split_by = "sex_final", 
                        show_midpoint = FALSE, 
                        pal = sexcolours) +
  
  # Update legend title:
  labs(fill = expr(bold("Geschlecht")), 
       x = "Altersklassen (Jahre)\n", 
       y = "\nAnzahl der Fälle") +
  
  # Hardcode y axis interval (number of cases) to 1:
  scale_y_continuous(breaks = integer_breaks()) +
  
  # Change theme to classic:
  theme_classic() +
  
  # Make axis titles bold:
  theme(axis.title = element_text(face = "bold", size = 10))

# Print the age sex pyramid:
agesexpyramid

```
<br>

```{r agesex_piebox, fig.height = 3.5, fig.cap = "(A) Boxplot des Alters; (B) Kuchendiagramm der Geschlechterverteilung"}

##########################
# Create age sex boxplot:

# Calculate median age by sex:
agesex_medians <- linelist %>% 
  
  group_by(sex_final) %>% 
  
  summarise(median = median(age_final, na.rm = TRUE))

# Extract overall median:
age_median <- compare %>% 
  
  filter(name == "median_age") %>% 
  
  select(value_current) %>% 
  
  pull()

# Create the boxplot:
age_boxplot <- linelist %>% 
  
  drop_na(age_final) %>% 
  
  ggplot() +
  
  # Define x and y axis and fill colours:
  geom_boxplot(mapping = aes(y = age_final), 
               fill = "#4F81BD") +
  
  # Limit to 100:
  coord_cartesian(ylim = c(0, 100)) +
  
  # Add classic theme (removes gridlines):
  theme_classic() +
  
  # Add axis labels and caption:
  labs(x = "Alle Fälle", 
       y = "Alter (Jahre)", 
       caption = stringr::str_glue("Medianalter insgesamt: {age_median} Jahre")) +
  
  # Format font of axis titles, text label and caption:
  theme(axis.title = element_text(face = "bold", size = 10),
        axis.text.y = element_text(size = 10), 
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        plot.caption = element_text(size = 10)) +
  
  # Add plot subtitle:
  ggtitle("A")


#################################
# Create sex distribution pie:

# Create table with percentages:
sexpie <- linelist %>%
  
  # Remove NAs:
  drop_na(sex_final) %>% 
  
  # Tabulate
  janitor::tabyl(sex_final) %>% 
  
  # Convert proportion to percentages:
  adorn_pct_formatting(digits = 0) %>% 
  
  # Add colours:
  mutate(hexcodes = sexcolours) %>% 
  
  # Calculate positions for the labels:
  mutate(csum = rev(cumsum(rev(n))), 
         pos = n/2 + lead(csum, 1),
         pos = if_else(is.na(pos), n/2, pos))


########################
# Create the pie chart:
sex_piechart <- ggplot(data = sexpie, 
                       mapping = aes(x = "" , 
                                     y = n, 
                                     fill = fct_inorder(sex_final))) +
  
  # Make a single column bar chart:
  geom_col(width = 1, color = 1) +
  
  # Convert to pie chart:
  coord_polar(theta = "y") +
  
  # Add the colours:
  scale_fill_manual(values = sexpie$hexcodes) +
  
  # Add the geom repel labels with percentages:
  ggrepel::geom_label_repel(data = sexpie,
                            aes(y = pos, label = percent),
                            size = 4.5, 
                            # Change text colour make visible with background:
                            color ="black",
                            nudge_x = 1, 
                            show.legend = FALSE) +
  
  # Add legend with text labels (short names):
  guides(fill = guide_legend(title = "Geschlecht")) +
  
  # Remove background grid:
  theme_void() +
  
  # Make legend title bold:
  theme(legend.title = element_text(size = 10, face = "bold")) +
  
  # Add subtitle with extra line breaks to align with taller boxplot:
  ggtitle("B\n\n\n\n")


#########################
# Print the two charts:

gridExtra::grid.arrange(
  age_boxplot,
  sex_piechart,
  ncol = 2)

```

<br>


```{r agesex_table, tab.cap = "Alters- und Geschlechtsverteilung"}

# Create summary table for age and sex:
agesextab <- linelist %>%
  
  # Select columns to include:
  select(age_final, sex_final) %>% 
  
  gtsummary::tbl_summary(
    
    # Specify column types to make sure correct statistic is applied:
    type = list(age_final ~ "continuous", 
                sex_final ~ "categorical"),
    
    # Calculate percentages from columns:
    percent = "column", 
    
    # Show missing values only if present:
    missing = "ifany", 
    
    # Text label for missing values:
    missing_text = "Unbekannt", 
    
    # Specify statistics:
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    
    # Make nice labels:
    label  = list(
      age_final ~ "Alter (Jahre)",                           
      sex_final ~ "Geschlecht")
    
  ) %>% 
  
  # Make variable names bold and italics:
  bold_labels() %>% 
  italicize_labels() %>% 
  
  # Modify header:
  modify_header(
    label = "**Charakteristisch**",
    stat_0 = "**Fälle insgesamt**\n_N_ = {N}",
    ) %>% 
  
  # Modify footnotes:
  modify_footnote(text_interpret = "md", 
                  all_stat_cols() ~ 
                    "Median (IQR) für Alter; n (%) für Geschlecht") %>%
  
  # Convert to flextable for printing:
  gtsummary::as_flex_table() %>% 
  
  # Fit columns to text:
  set_table_properties(layout = "autofit")

#######################
# Print table:

agesextab

```



\pagebreak


## Beschreibung der Krankheitsschwere


```{r disease_severity_table, tab.cap = "Krankenhausaufenthalte und Todesfälle"}

# Extract total number of cases with info on hospitalisation status:
n_hosp_info <- compare %>% 
  
  filter(name == "n_hosp_info") %>% 
  
  select(value_current) %>% 
  
  pull() %>%
  
  as.numeric()

# Extract total number of outbreak cases that died (any cause):
n_death_total <- compare %>% 
  
  filter(name %in% c("n_deaths_salmonella", "n_deaths_other")) %>% 
  
  select(value_current) %>% 
  
  pull() %>% 
  
  as.numeric() %>% 
  
  sum()

######################################################
# Create summary table for hospitalisation and deaths:
hospdeathtab <- linelist %>%
  
  # Select columns to include:
  select(hosp_final, death_final) %>% 
  
  # Convert not hospitalised (Nicth gestorben) factor level to NA:
  mutate(death_final = fct_recode(death_final,
                                  NULL = "Nicht gestorben")) %>% 
  
  # Create gtsummary table:
  gtsummary::tbl_summary(
    
    # Calculate percentages from columns:
    percent = "column", 
    
    # Show missing values only if present:
    missing = "no", 
    
    # Make nice labels:
    label  = list(
      hosp_final ~ "Fälle mit vollständigen Angaben zum Hospitalisierungsstatus", 
      death_final ~ "Todesfälle")
    
  ) %>% 
  
  # Add total non-missing for hospital and deaths:
  add_n(statistic = "{N_nonmiss}") %>% 
  
  # Make variable names bold and italics:
  bold_labels() %>% 
  italicize_labels() %>%
  
  # Modify header:
  modify_header(
    label = "**Charakteristisch**",
    stat_0 = "**Fälle**"
    ) %>% 
  
  # Modify footnotes:
  modify_footnote(text_interpret = "md", 
                  all_stat_cols() ~ 
                    "n (%) der Fälle mit vollständigen Angaben zum Hospitalisierungsstatus") %>%
  
  # Convert to flextable for printing:
  gtsummary::as_flex_table() %>% 
  
  # Make non-missing totals bold:
  flextable::bold(i = c(1,5), 
                  j = 2) %>% 
  
  # Replace N with correct hospitalised total:
  flextable::compose(i = 1, 
                     j = 2, 
                     as_paragraph(as_chunk(as.integer(n_hosp_info)))) %>% 
  
  # Replace N with correct deaths total:
  flextable::compose(i = 5, 
                     j = 2, 
                     as_paragraph(as_chunk(as.integer(n_death_total)))) %>% 
  
  # Replace NAs with 0 for deaths:
  flextable::compose(i = ~ stat_0 == "0 (NA%)",
                     j = 3,
                     as_paragraph(as_chunk("0 (0%)"))) %>% 
  
  # Set rows to colour scheme for hospital status:
  bg(i = 2, bg = "#002060") %>% 
  
  bg(i = 3, bg = "#00B0F0") %>% 
  
  bg(i = 4, bg = "#C2D69B") %>% 
  
  # Change colour of font for dark blue background to white:
  color(i = 2, color = "white") %>% 
  
  # Fit columns to text:
  set_table_properties(layout = "autofit")


#######################
# Print table:

hospdeathtab

```

<br>


```{r hospital_piechart, fig.height = 3.5, fig.cap = "Kuchendiagramm des Krankenhausaufenthaltsstatus"}

#################################
# Create table with percentages:
hospie <- linelist %>%
  
  # Remove NAs:
  drop_na(hosp_final) %>% 
  
  # Tabulate
  janitor::tabyl(hosp_final) %>% 
  
  # Convert proportion to percentages:
  adorn_pct_formatting(digits = 0) %>% 
  
  # Add short names for the labels:
  mutate(shortlabs = case_when(
    
    # Hospitalised due to Salmonella label:
    hosp_final == "Hospitalisierung aufgrund der Salmonellose" 
    ~ "Hospitalisiert aufgrund der Salmonellose",
    
    # Hospitalised due to other causes label:
    hosp_final == "Hospitalisierung aufgrund anderer Ursachen" 
    ~ "Hospitalisiert aufgrund anderer Ursachen",
    
    # Not hospitalised label:
    hosp_final == "Nicht hospitalisiert" 
    ~ "Nicht hospitalisiert"
    
    )) %>% 
  
  # Add colours:
  mutate(hexcodes = case_when(
    
    # Colour for salmonella hospitalisations:
    shortlabs == "Hospitalisiert aufgrund der Salmonellose" ~ "#002060",
    
    # Colour for other cause hospitalisations:
    shortlabs == "Hospitalisiert aufgrund anderer Ursachen" ~ "#00B0F0",
    
    # Colour for not hospitalised cases:
    shortlabs == "Nicht hospitalisiert" ~ "#C2D69B"
    
    )) %>% 
  
  # Calculate positions for the labels:
  mutate(csum = rev(cumsum(rev(n))), 
         pos = n/2 + lead(csum, 1),
         pos = if_else(is.na(pos), n/2, pos))


########################
# Create the pie chart:
piechart <- ggplot(data = hospie, 
                   mapping = aes(x = "" , 
                                 y = n, 
                                 fill = fct_inorder(shortlabs))) +
  
  # Make a single column bar chart:
  geom_col(width = 1, color = 1) +
  
  # Convert to pie chart:
  coord_polar(theta = "y") +
  
  # Add the colours:
  scale_fill_manual(values = hospie$hexcodes) +
  
  # Add the geom repel labels with percentages:
  ggrepel::geom_label_repel(data = hospie,
                            aes(y = pos, label = percent),
                            size = 4.5, 
                            # Change text colour make visible with background:
                            color = c("white", "white", "black"),
                            nudge_x = 1, 
                            show.legend = FALSE) +
  
  # Add legend with text labels (short names):
  guides(fill = guide_legend(title = "Fälle mit vollständigen Angaben zum Hospitalisierungsstatus")) +
  
  # Remove background grid:
  theme_void() +
  
  # Make legend title bold:
  theme(legend.title = element_text(size = 10, face = "bold"))


######################
# Print the piechart:  

piechart

```

<br>

```{r hospital_boxplot, fig.height = 3.5, fig.cap = "Boxplot des Alters nach Hospitalisierungsstatus"}

# Set hospital status colours:
hospcolours <- setNames(object = c("#002060", 
                                   "#00B0F0", 
                                   "#C2D69B"), 
                        nm = c('Salmonellose',
                               'Andere Ursache',
                               'Nicht hospitalisiert'))

# Define list of levels in hospital status data:
hospindata <- linelist %>% 
  
  select(hosp_labels) %>% 
  
  mutate(hosp_labels = as.character(hosp_labels)) %>% 
  
  drop_na() %>% 
  
  pull() %>% 
  
  unique()

# Trim hospcolours to levels in data:
hospcolours <- subset(hospcolours, names(hospcolours) %in% hospindata)

# Create boxplot:
hosp_boxplot <- linelist %>% 
  
  select(hosp_labels, age_final) %>% 

  drop_na() %>% 
  
  ggplot(mapping = aes(x = hosp_labels, 
                       y = age_final)) +
  
  geom_boxplot(outlier.colour = "darkred", 
               outlier.shape = 19,
               outlier.size = 3, 
               colour = "darkgrey", 
               fill = hospcolours, 
               drop = TRUE) +

  stat_summary(fun.y = median, 
               geom = "point", 
               shape = 23, 
               size = 4, 
               colour = "darkgrey", 
               fill = "white") +
  
  labs(x = "", 
       y = "Alter (Jahre)\n") +
  
  theme_classic() +
  
  theme(axis.title = element_text(face = "bold", size = 10))
  

##########################  
# Print boxplot:
hosp_boxplot

```

<br>

```{r severity_heatmap, tab.cap = "Hospitalisierungsstatus und Todesfälle nach Altersgruppe"}

# Create raw hospitalisations table:
hosptab <- linelist %>% 
  
  drop_na(hosp_final, age_cat) %>% 
  
  janitor::tabyl(hosp_final, age_cat) %>% 
  
  mutate(hosp_final = as.character(hosp_final)) %>% 
  
  mutate(hosp_final = case_when(
    
    hosp_final == "Hospitalisierung aufgrund der Salmonellose" 
    ~ "Krankenhausaufenthalt \n(Salmonellose)", 
    
    hosp_final == "Hospitalisierung aufgrund anderer Ursachen" 
    ~ "Krankenhausaufenthalt \n(andere Ursache)", 
    
    .default = hosp_final
    
  )) %>% 
  
  rename(Schweregrad = hosp_final)


# Create raw death table:
deathtab <- linelist %>% 
  
  drop_na(death_final, age_cat) %>% 
  
  janitor::tabyl(death_final, age_cat) %>% 
  
  mutate(death_final = as.character(death_final)) %>% 
  
  mutate(death_final = case_when(
    
    death_final == "Todesfälle aufgrund von Salmonellose" 
    ~ "Gestorben \n(Salmonellose)", 
    
    death_final == "Todesfälle durch andere Ursachen" 
    ~ "Gestorben \n(andere Ursache)",
    
    .default = death_final
    
  )) %>% 
  
  rename(Schweregrad = death_final) %>% 
  
  filter(Schweregrad != "Nicht gestorben")


# Bind the two tables together:
ageseveritytab <- bind_rows(hosptab, deathtab) 
  

# Convert to heatmap with custom function:
ageseverityprint <- heatmaptable(ageseveritytab) %>% 
  
  # Add an extra header row with label for age categories:
  add_header_row(values = c("", "Schweregrad", "Alterskategorien (Jahre)"), 
                 colwidths = c(1, 1, nlevels(linelist$age_cat))) %>%
  
  # Merge the other parts of the header row:
  merge_v(j = c(1:2), part = "header") %>% 
  
  # Make the column with the severity categories (2) bold:
  bold(j = 2) %>% 
  
  # Reduce font size for better fit:
  fontsize(size = 10, part = "all") %>% 
  
  # Set cell colours of row title column:
  bg(i = 3, j = 2, bg = "#C2D69B", part = "body") %>% 
  
  bg(i = 2, j = 2, bg = "#00B0F0", part = "body") %>%
  
  bg(i = 1, j = 2, bg = "#002060", part = "body") %>%
  
  # Set font colours of row title column:
  color(i = 1, j = 2, color = "white", part = "body") %>% 
  
  # Add internal horizontal borders to body:
  hline(j = 2:(nlevels(linelist$age_cat) + 2), 
        border = fp_border(), 
        part = "body") %>% 

  # Autofit to page:
  set_table_properties(layout = "autofit")

  
#################
# Print table:
ageseverityprint


```


\pagebreak


# Mikrobiologie

`r ftext("[Fügen Sie hier Zahlen und Text zur Mikrobiologie hinzu.]", prop = placeholder)`


\pagebreak


# Internationaler Zusammenhang


`r ftext("[Fügen Sie hier Informationen über den internationalen Kontext hinzu (falls relevant).]", prop = placeholder)`


\pagebreak


# Explorative Befragungen


`r ftext("[Fügen Sie hier Informationen über Falluntersuchungen und explorative Befragungen hinzu.]", prop = placeholder)`


\pagebreak


# Ansprechpartner


Dieser Lagebericht wurde von **`r params$author_name`** (`r params$author_affiliation`) verfasst. 

Für weitere Informationen wenden Sie sich bitte an die unten aufgeführten Ansprechpartner.

<br>

```{r contacts_table}

# Create a data.frame with the relevant information:
contacts <- data.frame(
  
  detail = c("Name", 
             "Abteilung",
             "Fachgebiet",
             "Adresse", 
             "Telefon", 
             "E-Mail"), 
  
  epi = c("Dr. Anika Meinen", 
          "Infektionsepidemiologie",
          "Gastrointestinale Infektionen, Zoonosen und tropische Infektionen", 
          "Robert Koch-Institut\n\nSeestraße 10\n\n13353 Berlin", 
          "030 18754 3663", 
          "[FG35-Salm@rki.de](mailto:FG35-Salm@rki.de)"), 
  
  lab = c("Dr. Sandra Simon", 
          "Infektionskrankheiten",
          "Bakterielle darmpathogene Erreger und Legionellen und Nationales Referenzzentrum für Salmonellen und andere bakterielle Enteritiserreger",
          "Robert Koch-Institut\n\nBurgstraße 37\n\n38855 Wernigerode", 
          "030 18754 4211", 
          "[NRZ-Salm-FG11@rki.de](mailto:NRZ-Salm-FG11@rki.de)")
  
) %>% 
  
  # Convert to flextable:
  qflextable() %>% 
  
  # Allow markdown formatting to hyperlink email addresses:
  ftExtra::colformat_md(.sep = "\n") %>% 
  
  # Remove all borders with a theme:
  theme_booktabs() %>% 
  
  # Make column titles bold:
  bold(part = "header") %>% 
  
  # Make details column bold:
  bold(j = ~detail) %>% 
  
  # Align text to top of cell:
  valign(valign = "top", part = "body") %>% 
  
  # Reduce line spacing:
  line_spacing(i = ~ detail == "Adresse", 
               space = 1, 
               part = "body") %>% 
  
  # Update header labels:
  set_header_labels(
    detail = "", 
    epi = "Ansprechpartner am FG35 RKI:", 
    lab = "Ansprechpartner am NRZ:"
  ) %>% 
  
  # Fit columns to text:
  set_table_properties(layout = "autofit")


# Print the table:
contacts

```

<br>

**Das nächste Update wird abhängig von der epidemiologischen Lage erscheinen.**

```{r remove_temp_data, results='hide'}

# Remove temporary example data if used:
if (params$example_data) {
  invisible(file.remove(tmp_current, tmp_previous))
}


```

